\sick, the spectroscopic inference crank

### 2014-08-23 document/manuscript.tex
emulateapj format

\documentclass{aastex}
\usepackage[utf8]{inputenc}

\newcommand{\sick}{\textit{sick}}
\newcommand{\article}{\textit{Article}}

@@

In this \article{} I present \sick{}: a well-tested, MIT-licensed probabilistic 
software package for inference from spectroscopic data. \sick{} employs an 
approximation to data-generating models. Instead of attempting to model all 
kinds of expensive astrophysical processes (e.g., supernova explosions, stars, 
and any other interesting astrophysical processes) at run-time, it performs 
efficient multi-dimensional interpolation of model spectra from pre-computed 
grids. This approach is suitable for a plethora of different astrophysical 
processes, allowing any user to easily specify a model using existing published 
spectral grids, and infer astrophysical properties from their data. Aspects of 
the probabilistic model are described in Section \ref{sec:model}. A number of 
examples are included in Sections \ref{sec:inference-test} and \ref{sec:examples}, 
where I present accurate and precise objective photospheric measurements using 
noisy spectra. I conclude in Section \ref{sec:conclusions} with references to 
the online documentation and applicability of the software.

@@

The first step of the (approximate) generative model is to interpolate a model 
spectrum. \sick{} employs the Quickhull algorithm \citep{quickhull} to linearly 
interpolate in multiple dimensions. Irregularly-spaced grids are perfectly 
acceptable; Quickhull does not require the grid to be rectangular or 
regularly-spaced points. Large grids of model spectra are cached for 
computational efficiency, allowing for the total size of the model grid to far 
exceed the available random access memory.\footnote{In practice this is performed 
by efficient memory-mapping.} I introduce an interpolation function $S(\bm{\psi})$ 
that produces a model spectrum with $J$ points $\{\lambda_j,I_j\}_{0}^{J}$ at 
any evaluable point $\bm{\psi} \equiv (T_{\rm eff},\log{g},{\rm [M/H]},[\alpha/{\rm Fe}])$ 
within the grid boundaries.

@@

We now have a \textit{generative model} for the data. The frequency (or 
probability) distribution ${p\left(F_i|\lambda_i,\sigma_i,\bm{\psi},z,\sigma_s,\{c\}_{0}^{m}\right)}$ 
for $F_i$ is given by

@@

Finally, I consider the handling of outliers in the data. These may be in the form 
of cosmic ray spikes, improper calibration of the data, telluric features, or simply 
poorly modelled spectral regions. Treatment of these artefacts is achieved using a 
Gaussian mixture model: a combination of two models. In the mixture model, the data 
are fit by the sum of amplitudes ($1 - P_o$ and $P_o$, respectively) of two 
distributions: the expected fluxes $E_i$, and a normal distribution centered along 
the continuum function $C_i$ with variance $s_{i}^2 + V_{o}$. This requires the 
inclusion of two (final) additional parameters: $P_o$ and $V_o$. The prior 
distribution function $p\left(V_o\right)$ requires $V_{o}$ to always be positive 
(Equation \ref{eq:default_priors}), and as such the outlier distribution will 
\textit{always} have a larger variance. Distributions of smaller variance are more 
informative, so conceptually a fit to the expected fluxes $E_{i}$ is generally 
preferred wherever possible. For brevity I define $\bm{\kappa} \equiv (\bm{\psi},\{z,\sigma_s,\{c_k\}_{k=0}^{m},f\}_{0}^{N_{c}})$, and the likelihood for the mixture model is given by

@@

With all of the combined effects there are 16 free parameters including $\bm{\psi}$, 
the astrophysical quantities of interest. For the examples presented in Section 
\ref{sec:examples}, this requires interpolation between 92 billion pixels in four 
dimensions. While the description might appear daunting, the problem is tractable, 
numerically efficient, and easy to configure. By default \sick{} numerically solves 
this problem in three sequential steps, which are described in the following 
sections: scattering, optimisation, and Monte-Carlo Markov Chain (MCMC) sampling.

@@

After the random scattering step, the most probable $\bm{\theta}$ point is used 
as an initial guess for numerical optimisation. A number of suitable minimisation 
algorithms are available in \sick{} through the SciPy \citep{scipy} optimization 
module. The Nelder-Mead algorithm \citep{nelder-mead} is employed by default, and 
the full suite of optimisation algorithms available is listed in the online 
documentation. Unlike other optimisation techniques, the Nelder-Mead algorithm 
does not approximate first- or second-order derivatives, and can consequently be 
less efficient than other approaches. However it is robust in high-dimensional 
space, even in the presence of substantial noise, and has been successfully 
employed in a wide range of engineering and scientific problems. The reader is 
encouraged to experiment with other optimisation algorithms if the Nelder-Mead 
algorithm proves unsuitable or takes an untenable amount of time. Since these 
optimisation algorithms are minimisation techniques and I seek to maximise the 
log-probability $\log{\left(\mathcal{P}\right)}$, I numerically optimise the 
parameters $\bm{\theta}$ by minimising the negative log-probability 
$-\log{\left(\mathcal{P}\right)}$.

@@

The random scattering and numerical optimisation steps efficiently provide an 
accurate estimate of the optimal parameters $\bm{\theta}$. Once the optimisation
 step is complete, \sick{} employs the affine-invariant ensemble sampler proposed 
 by \citet{goodman;weare}, and implemented by \citet{emcee}. The Metropolis-Hastings 
 MCMC algorithm is employed by default. \sick{} allows for the model settings to 
 be specified in a human-readable \textsc{yaml}- or \textsc{json}-formatted 
 configuration file\footnote{The reader is referred to the online documentation 
 for an example.}, where the number of Goodman \& Weare walkers can be specified, 
 as well as the number of samples to perform. When the optimisation step is used, 
 the initial points are taken from a small multi-dimensional ball around the 
 optimised parameters $\bm{\theta_{opt}}$. If the optimisation step is not 
 performed, initial values for the MCMC walkers are drawn in the same way as 
 described for the scattering process.

@@

Quantitatively ensuring numerical convergence for MCMC analyses remains an 
unsolved problem. There are a number of excellent resources on MCMC sampling 
which outline this issue in greater detail. The mean acceptance fraction, 
auto-correlation times, and values of the parameter chains themselves all 
provide reasonable proxy indicators of convergence. The user is encouraged to 
inspect these metrics and devise some heuristic for when convergence has 
unambiguously been achieved. The final state of every MCMC analysis is 
automatically saved by \sick{}. This allows users to resume their analysis from 
the most recent state if they believe the system has not converged.

@@

As an initial test of the probabilistic framework I have interpolated a 
noise-free synthetic spectrum which will act as a faux observation. A single 
channel from the AMBRE synthetic library \citep{ambre} has been employed. The 
spectra were calculated for the Gaia-ESO Survey \citep{gaia-eso} and range from 
475\,nm to 685\,nm. I have applied a number of transformations to the 
interpolated spectrum: a second-order polynomial enters multiplicatively to 
represent the continuum, the spectrum is redshifted, and fluxes are convolved 
to a spectral resolution of $\mathcal{R} \sim 10000$. The faux data are resampled 
to a uniform spacing of $\sim{}0.8$\,\AA{}, effectively discarding $99.5\%$ of 
the original pixels, which were sampled at $\sim4\times10^{-4}$\,nm. White noise  
has been added to replicate a S/N ratio of $\sim 7$ pixel$^{-1}$. I also assumed 
that the observed variance is systematically underestimated by 10 per cent: 
conceptually the faux data is $\sim{}$10 per cent noisier than what an observer 
would estimate from the flux counts.

Default priors and configurations were employed. 200 walkers were used to explore 
the parameter space for 1250 steps as burn-in, and another 250 steps were used to 
sample the posterior. The chains for each parameter are shown in Figure 
\ref{fig:chains}. At least by eye, the simulation appears to have converged after 
just 300 steps. The marginalised posterior distributions are shown in Figure 
\ref{fig:corner-inference}, with the accepted truth values marked. The true and 
inferred values for all parameters are tabulated in Table \ref{tab:inference-test}.

This test demonstrates that accurate and precise inference of stellar parameters 
can be made even in the presence of substantial noise. The true values of all 
parameters are recovered within the uncertainties, and the largest (relative) 
uncertainties are observed for continuum coefficients $b_1$ and $b_2$. Figure 
\ref{fig:corner-inference} shows there are strong covariances between these 
parameters. This is unsurprising given the shape of the spectrum (Figure 
\ref{fig:spectrum-inference}), which could reasonably be fit with a linear 
continuum, even though the continuum was truly modelled as a third order 
polynomial. The logarithm of additional fractional variance $\ln{f_{blue}}$ is 
precisely constrained, suggesting the noise model is excellently matched by the 
data. Given the simplicity of the noise model, the precision of inference on 
$\ln{f_{blue}}$ is unlikely to be recovered for real data.

\begin{figure}
\label{fig:chains}
\includegraphics[height=\textheight]{chains.pdf}
\caption{Points sampled by the 200 walkers at each step during the self-consistent 
inference test. The true values are marked in blue. The first 1250 steps are 
discarded as the burn-in period.}
\end{figure}

\begin{figure}
\label{fig:corner-inference}
\includegraphics[width=\textwidth]{corner.pdf}
\caption{Marginalised posterior distributions for all parameters $\bm{\theta}$ 
from a faux observation with spectral resolution $\mathcal{R} \sim 10,000$ and 
S/N ratio $\sim{}7$\,pixel$^{-1}$. True values are marked in blue. This figure 
demonstrates that precise inference of stellar parameters can be made with 
high-resolution spectra, even in the presence of substantial noise.}
\end{figure}

\begin{table}
\caption{True and inferred model parameters for the self-consistent inference test}
\label{tab:inference-test}
\begin{tabular}{llrr}
\hline
\hline
Parameter & Description & Truth & Inferred \\
\hline
$T_{\rm eff}$ & Effective photospheric temperature (K) & 5454 & 5483$_{-137}^{+125}$ \\
$\log{}g$ & Surface gravity & 4.124 & 4.084$_{-0.274}^{+0.262}$ \\
${\rm [Fe/H]}$  & Metallicity & --0.514 & $-0.512_{-0.160}^{+0.156}$ \\
$[\alpha/{\rm Fe}]$ & $\alpha$-element abundance & $+0.02$ & $+0.05_{-0.12}^{+0.13}$ \\
$v$     & Velocity (km s$^{-1}$) & 13.0 & $9.5_{-5.5}^{+5.3}$ \\
$\sigma_{blue}$ & Gaussian smoothing sigma (\AA{}) & 0.581 & 0.518$_{-0.34}^{+0.39}$ \\
$\ln{f_{blue}}$ & Logarithm of fractionally underestimated variance & $-2.30$ & $-2.28_{-0.02}^{+0.02}$ \\
$b_{0}$ & Continuum polynomial coefficient ($\times10^{-3}$) & 1.23 & $1.26_{-0.32}^{+0.32}$ \\
$b_{1}$ & Continuum polynomial coefficient & $-0.593$ & $-1.001_{-3.634}^{+3.681}$ \\
$b_{2}$ & Continuum polynomial coefficient & $-0.569$ & $894_{-10335}^{+10367}$ \\
\hline
\end{tabular}
\end{table}

\begin{figure}
\label{fig:spectrum-inference}
\includegraphics[width=\textwidth]{spectrum.pdf}
\caption{A faux observed spectrum (black) of $\mathcal{R} \sim 10000$ and S/N 
ratio of $\sim7$ with an underestimated variance by $\sim$10\%. The maximum 
likelihood model spectrum is shown in red. }
\end{figure}

@@

\subsection{Sol}
\label{sec:sun}
A high-resolution ($\mathcal{R} \sim$ 20000), high S/N ($\sim{}$150 pixel$^{-1}$) 
twilight spectrum was obtained from the GIRAFFE/FLAMES archive
\footnote{https://www.eso.org/observing/dfo/quality/GIRAFFE/pipeline/solar.html}. 
The H875.7 setup was chosen for this example, yielding spectra from 848.2\,nm to 
899.2\,nm. Data points redder than 874\,nm have been discarded to more closely 
match the Gaia RVS wavelength coverage, and demonstrate the recoverability of 
stellar parameters on real data.

The AMBRE \citep{ambre} synthetic grid was also employed for this example. The 
continuum was represented with a third order polynomial, and no redshift or 
outlier modelling was included. The central $\pm$0.1 nm cores of the Ca II 
near-infrared triplet lines were masked (specifically rest wavelengths 
849.7-849.9 nm, 854.1-854.3, and 866.1-866.3 nm), as the cores are strongly 
affected by non-LTE effects that are not treated in the 1D LTE model atmospheres 
used to generate the AMBRE spectra. Pixels between rest wavelengths 850.10\,nm 
to 850.37\,nm were also masked as the models did not seem to accurately reproduce 
the data in this region.

\begin{figure}
\label{fig:solar}
\includegraphics[width=\textwidth]{solar.pdf}
\caption{Stellar parameter ($\bm{\psi}$: $T_{\rm eff}$, $\log{g}$, [Fe/H], 
[$\alpha$/Fe])  posterior distributions for the GIRAFFE/FLAMES twilight spectrum. 
Nuisance parameters are not shown. Maximum likelihood values for each parameter 
are drawn in red, and the 16\%, 50\% and 84\% quantiles are shown as dashed lines. 
The posterior distributions are precise, and close to the accepted values (not 
shown) of $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, slightly discrepant from 
[$\alpha$/Fe] = 0, and significantly deviant from [Fe/H] = 0. This test 
illustrates the precision achievable using \sick, but highlights the need for 
accurate models.}
\end{figure}

For this example 1000 draws were made during the random scattering stage, and 
optimisation was performed using the default options. One hundred walkers 
explored the parameter space for 1000 steps, before another 200 steps were used 
to sample the posterior. The posterior distributions for stellar parameters is 
shown in Figure \ref{fig:solar}. The currently accepted stellar parameters for 
the Sun are $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, [Fe/H] = 0, and 
[$\alpha$/Fe] = 0. The inferred parameters agree excellently with the accepted 
values for effective temperature and surface gravity: 
$T_{{\rm eff},inferred} = 5770^{+20}_{-14}$, and $\log{g}_{inferred} = 4.40^{+0.06}_{-0.05}$. 
The agreement is less pleasing for $\alpha$-element enhancement, but less so for 
mean metallicity: [$\alpha$/Fe]$_{inferred} = -0.03^{+0.01}_{-0.01}$, and 
[Fe/H]$_{inferred} = -0.07^{+0.01}_{-0.01}$.

While our inferences are precise, the offset in mean metallicity suggest the 
models may be slightly inaccurate in this wavelength regime or that there are 
unaccounted telluric features in the data. This test highlights the need to 
verify the accuracy of the input models. A useful extension of this project 
would be to simultaneously include telluric spectra, as well as pre-computed 
astereoseismic models, such that spectroscopy and time-series photometry could 
simultaneously constrain stellar parameters. Asteroseismology can accurately and 
precisely constrain surface gravity, but requires an input effective temperature. 
Conversely, surface gravity is the least constrained parameter by spectroscopy, 
but can precisely constrain effective temperature is the only method available 
to measure chemical abundances. Since \sick{} allows for any dispersion unit 
equivalency (e.g., frequency, wavelength, etc), spectroscopic and astereoseismic 
models can be trivially coupled into a generative framework for unprecedented 
accuracy and precision in stellar spectroscopy.

\subsection{Globular Clusters}
Globular cluster stars are excellent candidates for testing \sick{} over a large 
range of stellar parameters. Here I consider a realistic scenario of 
low-resolution, noisy spectra of red giant and asymptotic branch cluster 
candidates observed using the AAOmega spectrograph on the Anglo-Australian 
Telescope in Coonabarabran, Australia. The 580V and 1700D gratings were employed, 
yielding one channel between 370\,nm to 580\,nm with a spectral resolution 
$\mathcal{R} \sim 1300$, and another from 820\,nm to 890\,nm with spectral 
resolution $\mathcal{R} \sim 8000$.

@@

Radial velocities were used to identify and discard non-cluster members. The 
metallicity distribution of the remaining cluster stars are shown in Figure 
\ref{fig:clusters}, where a comparison to the literature is made. The 
\citet{harris} catalogue (accessed June 2014) has been used as a literature 
reference for all globular clusters. Note that while there are measurable 
intrinsic metallicity spreads for these clusters, I have not adopted these as 
x-axis uncertainties.

\begin{figure}
\label{fig:clusters}
\includegraphics[width=\textwidth]{clusters.pdf}
\caption{Metallicities of cluster stars inferred from noisy, low-resolution 
spectra obtained with the AAOmega spectrograph. The agreement with the 
literature values (see text) is very reasonable.}
\end{figure}

Even ignoring uncertainties in literature values, the agreement between is 
exceptional for noisy, low-resolution spectra. No systematic trend is present 
and a negligible mean metallicity difference of $-0.001$ dex (accuracy of 0.2\%) 
is observed over three orders of magnitude. The agreement would be remarkable 
for high-resolution spectra with exquisite S/N ratios. I emphasise that these 
results were obtained without performing any `calibration' or ad-hoc posterior 
corrections, which are usually entirely empirical and not motivated by an 
astrophysical understanding.

Individual uncertainties in stellar parameters are also extremely reasonable, 
particularly given the low spectral resolution and S/N of the data. Uncertainties 
of $\sim$75 K in $T_{\rm eff}$, $\sim0.2$ in $\log{g}$, $\sim$0.1 dex in [Fe/H] 
and $\sim0.05$ dex in [$\alpha$/Fe] are obtained in the presence of substantial 
noise (e.g., $S/N \sim 15$). The most uncertain results were obtained for 
NGC 7089, where the lowest S/N ratios exist and only 6 cluster members were 
observed.

>>>>

\documentclass[iop]{emulateapj}
\usepackage[utf8]{inputenc}

\newcommand{\sick}{\texttt{sick}}
\newcommand{\article}{\textit{Article}}

@@

In this \article{} I present \sick{}: a well-tested, MIT-licensed probabilistic 
software package for inference from spectroscopic data. \sick{} employs an 
approximation to data-generating models. Instead of attempting to model all 
kinds of expensive astrophysical processes (e.g., stars, supernova, 
and any other interesting astrophysical processes) at run-time, it performs 
efficient multi-dimensional interpolation of model spectra from pre-computed 
grids and includes contributory phenomena (e.g., continuum, redshift) as free
parameters within an objective scalar-justified model. This approach is suitable 
for a plethora of different astrophysical 
processes, allowing any user to easily specify a model using existing published 
spectral grids, and infer astrophysical properties from their data. Aspects of 
the probabilistic model are described in Section \ref{sec:model}. A number of 
examples are included in Sections \ref{sec:inference-test} and \ref{sec:examples}, 
where I present accurate and precise objective photospheric measurements using 
noisy spectra. I conclude in Section \ref{sec:conclusions} with references to 
the online documentation and applicability of the software.

@@

The first step of the (approximate\footnote{The term approximate is used because 
model spectra are interpolated.}) generative model is to interpolate a model 
spectrum. \sick{} employs the Quickhull algorithm \citep{quickhull} to linearly 
interpolate in multiple dimensions. Irregularly-spaced grids are perfectly 
acceptable; Quickhull does not require the grid to be rectangular or 
regularly-spaced points. Large grids of model spectra are cached for 
computational efficiency, allowing for the total size of the model grid to far 
exceed the available random access memory.\footnote{In practice this is performed 
by efficient memory-mapping.} I introduce an interpolation function $S(\bm{\psi})$ 
that produces a model spectrum with $J$ points $\{\lambda_j,I_j\}_{0}^{J}$ at 
any evaluable point $\bm{\psi} \equiv (T_{\rm eff},\log{g},{\rm [M/H]},[\alpha/{\rm Fe}])$ 
within the grid boundaries.

@@

We now have a \textit{generative model} for the data. The frequency (or 
probability) distribution ${p\left(F_i|\lambda_i,\sigma_i,\bm{\psi},z,\sigma_s,\{c\}_{0}^{m}\right)}$ 
for the observed data $F_i$ is given by

@@

Finally, I consider the handling of outliers in the data. These may be in the form 
of cosmic ray spikes, improper calibration of the data, telluric features, or simply 
poorly modelled spectral regions. Treatment of these artefacts is achieved using a 
Gaussian mixture model: a combination of two models. In a mixture model, the data 
are fit by the sum of amplitudes ($1 - P_o$ and $P_o$, respectively) of two 
distributions: the expected fluxes $E_i$, and a normal distribution centered along 
the continuum function $C_i$ with variance $s_{i}^2 + V_{o}$. This requires the 
inclusion of two  additional parameters: $P_o$ and $V_o$. The prior 
distribution function $p\left(V_o\right)$ requires $V_{o}$ to always be positive 
(Equation \ref{eq:default_priors}), and as such the outlier distribution will 
\textit{always} have a larger variance. Distributions of smaller variance are more 
informative, so conceptually a fit to the expected fluxes $E_{i}$ is generally 
preferred wherever possible. For brevity I define $\bm{\kappa} \equiv (\bm{\psi},\{z,\sigma_s,\{c_k\}_{k=0}^{m},f\}_{0}^{N_{c}})$, and the likelihood for the mixture model is given by

@@

With all of the combined effects there are 16 free parameters including $\bm{\psi}$, 
the astrophysical quantities of interest. For the examples presented in Section 
\ref{sec:examples}, this requires interpolation between $9.2 \times 10^{10}$ pixels in four 
dimensions. While the description might appear daunting, the problem is tractable, 
numerically efficient, and easy to configure. By default \sick{} numerically solves 
this problem in three sequential steps, which are described in the following 
sections: scattering, optimisation, and Monte-Carlo Markov Chain (MCMC) sampling.

@@

After the random scattering step, the most probable $\bm{\theta}$ point is used 
as an initial guess for numerical optimisation. A number of suitable minimisation 
algorithms are available in \sick{} through the SciPy \citep{scipy} optimization 
module. The \citet{nelder-mead} algorithm is employed by default.
Unlike other optimisation techniques, the Nelder-Mead algorithm 
does not approximate first- or second-order derivatives, and can consequently be 
less efficient than other approaches. However it is robust in high-dimensional 
space, even in the presence of substantial noise, and has been successfully 
employed in a wide range of engineering and scientific problems. The reader is 
encouraged to experiment with other optimisation approaches if the Nelder-Mead 
algorithm proves unsuitable or takes an untenable amount of time. Since these 
optimisation algorithms are minimisation techniques and I seek to maximise the 
log-probability $\log{\left(\mathcal{P}\right)}$, I numerically optimise the 
parameters $\bm{\theta}$ by minimising the negative log-probability 
$-\log{\left(\mathcal{P}\right)}$.

@@

The random scattering and numerical optimisation steps efficiently provide an 
accurate estimate of the optimal parameters $\bm{\theta_{opt}}$. Once the optimisation
 step is complete, \sick{} employs the affine-invariant ensemble sampler proposed 
 by \citet{goodman;weare}, and implemented by \citet{emcee}. The Metropolis-Hastings 
 MCMC algorithm is employed by default. \sick{} allows for the model settings to 
 be specified in a human-readable \textsc{yaml}- or \textsc{json}-formatted 
 configuration file\footnote{The reader is referred to the online documentation 
 for an example.}, where the number of Goodman \& Weare walkers can be specified, 
 as well as the number of samples to perform. When the optimisation step is used, 
 the initial points are taken from a small multi-dimensional ball around the 
 optimised parameters $\bm{\theta_{opt}}$. If the optimisation step is not 
 performed, initial values for the MCMC walkers are drawn in the same way as 
 described for the scattering process.

@@

Quantitatively ensuring numerical convergence for MCMC analyses remains an 
unsolved problem. There are a number of excellent resources on MCMC sampling 
which outline this issue in greater detail. The mean acceptance fraction, 
auto-correlation times, and values of the parameter chains themselves all 
provide reasonable proxy indicators of convergence. The user is encouraged to 
inspect these metrics and devise some heuristic for when convergence has 
unambiguously been achieved. The final state of every MCMC analysis is 
automatically saved by \sick{}, allowing users to resume their analysis from 
the most recent state if they believe the system has not converged.

@@

As an initial test of the probabilistic framework I have interpolated a 
noise-free synthetic spectrum which will act as a faux observation. A single 
channel from the AMBRE synthetic library \citep{ambre} has been employed. These 
spectral channels range from 475\,nm to 685\,nm and were specifically calculated 
by the AMBRE group for the Gaia-ESO Survey \citep{gaia-eso}. I have applied a number of transformations to the 
interpolated spectrum: a second-order polynomial enters multiplicatively to 
represent the continuum, the spectrum is redshifted, and fluxes are convolved 
to a spectral resolution of $\mathcal{R} \sim 10000$. The faux data are resampled 
to a uniform spacing of $\sim{}0.08$\,nm, effectively discarding $99.5\%$ of 
the original pixels, which were sampled at $\sim4\times10^{-4}$\,nm. White noise  
has been added to replicate a S/N ratio of $\sim 7$ pixel$^{-1}$. I also assumed 
that the observed variance is systematically underestimated by 10 per cent: 
conceptually the faux data is $\sim{}$10 per cent noisier than what an observer 
would estimate from the flux counts.

\begin{figure*}
\label{fig:chains}
\includegraphics[height=\textheight]{chains.pdf}
\caption{Points sampled by the 200 walkers at each step during the self-consistent 
inference test. The true values are marked in blue. The first 1250 steps are 
discarded as the burn-in period.}
\end{figure*}

\begin{figure*}
\label{fig:corner-inference}
\includegraphics[width=\textwidth,height=\textwidth]{corner.pdf}
\caption{Marginalised posterior distributions for all parameters $\bm{\theta}$ 
from a faux observation with spectral resolution $\mathcal{R} \sim 10,000$ and 
S/N ratio $\sim{}7$\,pixel$^{-1}$. True values are marked in blue. This figure 
demonstrates that precise inference of stellar parameters can be made with 
high-resolution spectra, even in the presence of substantial noise.}
\end{figure*}

\begin{table*}
\center
\caption{True and inferred model parameters for the self-consistent inference test}
\label{tab:inference-test}
\begin{tabular}{llrr}
\hline
\hline
Parameter & Description & Truth & Inferred \\
\hline
$T_{\rm eff}$ & Effective photospheric temperature (K) & 5454 & 5483$_{-137}^{+125}$ \\
$\log{}g$ & Surface gravity & 4.124 & 4.084$_{-0.274}^{+0.262}$ \\
${\rm [Fe/H]}$  & Metallicity & --0.514 & $-0.512_{-0.160}^{+0.156}$ \\
$[\alpha/{\rm Fe}]$ & $\alpha$-element abundance & $+0.02$ & $+0.05_{-0.12}^{+0.13}$ \\
$v$     & Velocity (km s$^{-1}$) & 13.0 & $9.5_{-5.5}^{+5.3}$ \\
$\sigma_{blue}$ & Gaussian smoothing sigma (\AA{}) & 0.581 & 0.518$_{-0.34}^{+0.39}$ \\
$\ln{f_{blue}}$ & Logarithm of fractionally underestimated variance & $-2.30$ & $-2.28_{-0.02}^{+0.02}$ \\
$b_{0}$ & Continuum polynomial coefficient ($\times10^{-3}$) & 1.23 & $1.26_{-0.32}^{+0.32}$ \\
$b_{1}$ & Continuum polynomial coefficient & $-0.593$ & $-1.001_{-3.634}^{+3.681}$ \\
$b_{2}$ & Continuum polynomial coefficient & $-0.569$ & $894_{-10335}^{+10367}$ \\
\hline
\end{tabular}
\end{table*}

\begin{figure*}
\label{fig:spectrum-inference}
\includegraphics[width=\textwidth]{spectrum.pdf}
\caption{A faux observed spectrum (black) of $\mathcal{R} \sim 10000$ and S/N 
ratio of $\sim7$ with an underestimated variance by $\sim$10\% which was used for 
the self-consistent inference test. The recovered maximum likelihood model spectrum 
is shown in red. }
\end{figure*}

Default priors and configurations were employed. 200 walkers were used to explore 
the parameter space for 1250 steps as burn-in, and another 250 steps were used to 
sample the posterior. The chains for each parameter are shown in Figure 
\ref{fig:chains}. At least by eye, the chains appears to have converged after 
just 300 steps. The marginalised posterior distributions are shown in Figure 
\ref{fig:corner-inference}, with the accepted truth values marked. The true and 
inferred values for all parameters are tabulated in Table \ref{tab:inference-test}.

This test demonstrates that accurate and precise inference of stellar parameters 
can be made even in the presence of substantial noise. The true values of all 
parameters are recovered within the uncertainties, and the largest (relative) 
uncertainties are observed for continuum coefficients $b_1$ and $b_2$. Figure 
\ref{fig:corner-inference} shows there are strong covariances between these 
parameters. This is unsurprising given the shape of the spectrum (Figure 
\ref{fig:spectrum-inference}), which could reasonably be fit with a linear 
continuum, even though the continuum was truly represented by a third order 
polynomial. The logarithm of additional fractional variance $\ln{f_{blue}}$ is 
precisely constrained, suggesting the noise model is excellently matched by the 
data. Given the simplicity of the noise model, the precision of inference on 
$\ln{f_{blue}}$ is unlikely to be recovered for real data.

@@

\subsection{Sol}
\label{sec:sun}
A high-resolution ($\mathcal{R} \sim$ 20000), high S/N ($\sim{}$150 pixel$^{-1}$) 
twilight spectrum was obtained from the GIRAFFE/FLAMES 
archive\footnote{eso.org/observing/dfo/quality/GIRAFFE/pipeline/solar.html}. 
The H875.7 setup was chosen for this example, yielding spectra from 848.2\,nm to 
899.2\,nm. Data points redder than 874\,nm have been discarded to more closely 
match the Gaia RVS wavelength coverage, and demonstrate the recoverability of 
stellar parameters on real data.

The AMBRE \citep{ambre} synthetic grid was also employed for this example. The 
continuum was represented with a third order polynomial, and no redshift or 
outlier modelling was included. The central $\pm$0.1 nm cores of the Ca II 
near-infrared triplet lines were masked (specifically rest wavelengths 
849.7-849.9\,nm, 854.1-854.3\,nm, and 866.1-866.3\,nm), as the cores are strongly 
affected by non-LTE effects that are not accounted for in the 1D LTE model atmospheres 
used to generate the AMBRE spectra. Pixels between rest wavelengths 850.10\,nm 
to 850.37\,nm were also masked as the models did not seem to accurately reproduce 
the data in this region.

\begin{figure*}
\label{fig:solar}
\includegraphics[width=\textwidth,height=\textwidth]{solar.pdf}
\caption{Stellar parameter ($\bm{\psi}$: $T_{\rm eff}$, $\log{g}$, [Fe/H], 
[$\alpha$/Fe])  posterior distributions for the GIRAFFE/FLAMES twilight spectrum. 
Nuisance parameters are not shown. Maximum likelihood values for each parameter 
are drawn in red, and the 16\%, 50\% and 84\% quantiles are shown as dashed lines. 
The posterior distributions are precise, and close to the accepted values (not 
shown) of $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, slightly discrepant from 
[$\alpha$/Fe] = 0, and significantly deviant from [Fe/H] = 0. This test 
illustrates the precision achievable using \sick, but highlights the need for 
accurate models.}
\end{figure*}

For this example 1000 draws were made during the random scattering stage, and 
optimisation was performed using the default options. One hundred walkers 
explored the parameter space for 1000 steps, before another 200 steps were used 
to sample the posterior. The posterior distributions for stellar parameters is 
shown in Figure \ref{fig:solar}. The currently accepted stellar parameters for 
the Sun are $T_{\rm eff} = 5777$\,K, $\log{g} = 4.445$, [Fe/H] = 0, and 
[$\alpha$/Fe] = 0. The inferred parameters agree excellently with the accepted 
values for effective temperature and surface gravity: 
$T_{{\rm eff},inferred} = 5770^{+20}_{-14}$, and $\log{g}_{inferred} = 4.40^{+0.06}_{-0.05}$. 
The agreement is less pleasing for $\alpha$-element enhancement, and worse for 
mean metallicity: [$\alpha$/Fe]$_{inferred} = -0.03^{+0.01}_{-0.01}$, and 
[Fe/H]$_{inferred} = -0.07^{+0.01}_{-0.01}$.

While our inferences are precise, the offset in mean metallicity suggest the 
models may be slightly inaccurate in this wavelength regime, or the solar abundances
employed are slightly discrepant, or perhaps there are 
unaccounted telluric features in the data. This test highlights the need to 
verify the accuracy of the input models. A useful extension of this project 
would be to simultaneously include telluric spectra, as well as pre-computed 
astereoseismic models, such that spectroscopy and time-series photometry could 
simultaneously constrain stellar parameters. Asteroseismology can accurately and 
precisely constrain surface gravity, but requires an input effective temperature. 
Conversely surface gravity is the least constrained stellar parameter by spectroscopy, 
but can precisely constrain effective temperature, and is the only method available 
to measure chemical abundances. Since \sick{} allows for any dispersion unit 
equivalency (e.g., frequency, wavelength, et cetera), spectroscopic and astereoseismic 
models can be trivially coupled into a generative framework for unprecedented 
accuracy and precision in stellar spectroscopy.

\subsection{Globular Clusters}
Globular cluster stars are excellent candidates for testing \sick{} over a large 
range of model parameters. Here I consider a realistic scenario of 
low-resolution, noisy spectra of red giant and asymptotic branch cluster 
candidates observed using the AAOmega spectrograph on the Anglo-Australian 
Telescope in Coonabarabran, Australia. The 580V and 1700D gratings were employed, 
yielding one channel between 370\,nm to 580\,nm with a spectral resolution 
$\mathcal{R} \sim 1300$, and another from 820\,nm to 890\,nm with spectral 
resolution $\mathcal{R} \sim 8000$.

@@

Radial velocities were used to identify and discard non-cluster members. The 
metallicity distribution of the remaining cluster stars are shown in Figure 
\ref{fig:clusters}, where a comparison to the literature is made. The 
\citet{harris} catalogue (accessed June 2014) has been used as a literature 
reference for all globular clusters. Note that although there are measurable 
intrinsic metallicity spreads and active debates on the mean abundances 
of these clusters, I have not adopted abscissa uncertainties.

\begin{figure}[h!]
\label{fig:clusters}
\includegraphics[width=0.5\textwidth]{clusters.pdf}
\caption{Metallicities of cluster stars inferred from noisy, low-resolution 
spectra obtained with the AAOmega spectrograph. The agreement with the 
literature values (see text) is very reasonable.}
\end{figure}

Even ignoring uncertainties in the literature values, the agreement between is 
exceptional for noisy, low-resolution spectra. No systematic trend is present 
and a negligible mean metallicity difference of $-0.001$ dex (an accuracy of $<$0.2\%) 
is observed over three orders of magnitude. The agreement would be remarkable 
for high-resolution spectra with exquisite S/N ratios. I emphasise that these 
results were obtained without performing any `calibration' or ad-hoc posterior 
corrections, which are usually entirely empirical and not motivated by an 
astrophysical understanding.

Individual uncertainties in stellar parameters are also extremely reasonable, 
particularly given the low spectral resolution and S/N of the data. Uncertainties 
of $\sim$75 K in $T_{\rm eff}$, $\sim$0.2 in $\log{g}$, $\sim$0.1 dex in [Fe/H] 
and $\sim$0.05 dex in [$\alpha$/Fe] are obtained in the presence of substantial 
noise (e.g., $S/N \sim 15$). The most uncertain results were obtained for 
{NGC\,7089}, where the lowest S/N ratios were achieved and only six cluster members were 
observed.
